---
title: "Classifiers NEW"
author: "Ani Khachatryan"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='E:/Columbia/Spring2020/lab/data/')
```

Global options
```{r}
BALANCED_TEST <- TRUE
# TRUE because FALSE leads to horrible performance
BALANCED_TRAIN <- TRUE
ORGAN_SPECIFIC <- TRUE
CONVERT_TO_RANK_PERC <- TRUE
CROSS_VALIDATE <- TRUE
k <- 3
```

Imports
```{r}
# LASSO
library(glmnet)
library(tidyverse)
library(dplyr)

# Random Forest
library(randomForest)
require(caTools)

# Gradient Boosting
require(gbm)

# cross-validation
library(caret)

# roc curves
library(pROC)
```


Load the risk genes.
```{r}
DDD_all <- read.csv('./DDD/DDG2P_28_1_2020.csv')

DDD_all$organ.specificity.list <- tolower(DDD_all$organ.specificity.list)
DDD_all$gene.symbol <- tolower(DDD_all$gene.symbol)

DDD <- subset(DDD_all, DDD.category == 'confirmed')
```

Create a T/F table of genes vs RBPs, TRUE if RBP regulates the gene
rbp_threshold is the number of binding sites/events (?) necessary to consider the RBP a regulator of the protein

```{r}
# RBP-regualted genes
rbp_table <- read.table('./RBP_targets/RBP_targets.txt', header=TRUE, row.names = 1)
colnames(rbp_table) <- tolower(colnames(rbp_table))
rownames(rbp_table) <- tolower(rownames(rbp_table))
rbps_w_celltypes <- colnames(rbp_table)

sfun <- function(MyString) {
  return(strsplit(MyString, '.', TRUE)[[1]][2])
}

rbps <- sapply(rbps_w_celltypes, sfun)
#rbps <- tolower(sapply(rbps_w_celltypes, sfun))
colnames(rbp_table) <- rbps

# needs to have at least rbp_threshold binding sites 
rbp_threshold <- 10
rbp_table_tf <- rbp_table >= rbp_threshold
```

Helper function: removes the extension from the cell type
```{r}
remove_ext <- function(mystr) {
    return(tools::file_path_sans_ext(mystr))
  }
```

Helper function: replaces spaces in a string with underscores
```{r}
replace_space <- function(mystr) {
  return(gsub(' ', '_', mystr, fixed = TRUE))
}
```

Helper function: adds axtension
```{r}
add_ext <- function(mystr, myext) {
    return(paste0(mystr, '.', myext))
  }
```

Helper function: takes a vector of values and converts them to rank percentile
```{r}
rank_percentile <- function(mydata) {
  mydata.rank <- rank(mydata, ties.method="min")
  mydata.rank.perc <- round(mydata.rank/length(mydata.rank), 4)
  
  return(mydata.rank.perc)
}
```

A function that takes the data as a data.frame and returns a data frame that can be used as an input to the classifier
gene_label is the label - risk/non-risk
```{r}
make_classifier_df <- function(data_df, organ=NULL, convert_to_rank_perc=FALSE, expr_only=FALSE) {
  risk_genes <- DDD$gene.symbol
  if (!is.null(organ)) {
    risk_genes <- unique(DDD[grep(organ, DDD$organ.specificity.list, ignore.case = TRUE),]$gene.symbol)
  }
  
  rownames(data_df) <- tolower(rownames(data_df))
  colnames(data_df) <- unlist(lapply(colnames(data_df), replace_space))
  colnames(data_df) <- unlist(lapply(colnames(data_df), remove_ext))
  colnames(data_df) <- unlist(lapply(colnames(data_df), add_ext, myext='ct'))


  # only keep genes for which we have both rbp and expression data
  data_df <- data_df[rownames(data_df) %in% rownames(rbp_table_tf),]
  
  if (!expr_only) {
    rbp_table_tf_cur <- rbp_table_tf[rownames(rbp_table_tf) %in% rownames(data_df),]
    # make colnames unique 
    colnames(rbp_table_tf_cur) <- make.unique(colnames(rbp_table_tf_cur))
    # label genes risk or non-risk, remove possible risk genes
    rbp_table_tf_cur <- rbp_table_tf_cur[((rownames(rbp_table_tf_cur) %in% risk_genes) + (rownames(rbp_table_tf_cur) %in% DDD_all$gene.symbol) != 1),]
    colnames(rbp_table_tf_cur) <- unlist(lapply(colnames(rbp_table_tf_cur), add_ext, myext='rbp'))
  }

  data_df <- data_df[((rownames(data_df) %in% risk_genes) + (rownames(data_df) %in% DDD_all$gene.symbol) != 1),]
  
  # label
  gene_label <- rownames(data_df) %in% risk_genes
  gene_label <- ifelse(gene_label, 1, 0)

  # aggregate expression by celltype
  # same as data_df but transposed, genes are columns and samples are rows now
  data_df_t <- t(data_df)
  
  # genes are columns and rows are aggregate samples
  data_df_t_agg <- aggregate.data.frame(x = data_df_t, by = list(row.names(data_df_t)), FUN = mean)
  rownames(data_df_t_agg) <- data_df_t_agg$Group.1
  data_df_agg <- t(data_df_t_agg[,-1])
  
  # convert expression values to rank percentile
  if (convert_to_rank_perc) {
    data_df_agg <- apply(X=data_df_agg, MARGIN = 2, FUN = rank_percentile)
  }
  if (!expr_only) {
  # cbind rbp and expression data
    classifier_df <- cbind(rbp_table_tf_cur, data_df_agg)
  } else {
    classifier_df <- data_df_agg
  }
  classifier_df <- cbind(classifier_df, gene_label)

  return(as.data.frame(classifier_df))
}
```

Takes the input data and splits it into training and testing sets
```{r}
create_train_test_data <- function(inputData, balanced_test=FALSE, balanced_train=FALSE, train_proportion=.7) {
  # Create Training Data
  input_ones <- inputData[which(inputData$gene_label == 1), ]  # all 1's
  input_zeros <- inputData[which(inputData$gene_label == 0), ]  # all 0's
  input_ones_training_rows <- sample(1:nrow(input_ones), train_proportion*nrow(input_ones))  # 1's for training
  if (balanced_train) {
    # 0's for training. Pick as many 0's as 1's
    input_zeros_training_rows <- sample(1:nrow(input_zeros), train_proportion*nrow(input_ones)) 
  } else {
    # pick more 0s than ones, proportionally to the data
    input_zeros_training_rows <- sample(1:nrow(input_zeros), train_proportion*nrow(input_zeros))
  }

  training_ones <- input_ones[input_ones_training_rows, ]  
  training_zeros <- input_zeros[input_zeros_training_rows, ]
  trainingData <- rbind(training_ones, training_zeros)  # row bind the 1's and 0's 
  
  # Create Test Data
  test_ones <- input_ones[-input_ones_training_rows, ]
  test_zeros <- input_zeros[-input_zeros_training_rows, ]
  if (balanced_test) {
    test_zeros <- sample_n(test_zeros, dim(test_ones)[1])
  }
  
  testData <- rbind(test_ones, test_zeros)  # row bind the 1's and 0's 
  
  return(list(trainingData, testData))
}
```

**Logistic Regression**
```{r warning=FALSE}
do_logistic <- function(trainingData, testData, pval_threshold = .1, return_probs=FALSE, cv=CROSS_VALIDATE, k=3) {
  if (cv) {
    X <- rbind(trainingData[, -which(names(trainingData) == "gene_label")],  testData[, -which(names(testData) == "gene_label")])
    Y <- as.factor(c(trainingData$gene_label, testData$gene_label))
    train.control <- trainControl(method = "cv", number=k)
    model.fit.cv <- train(X, Y, method = "glm", family="binomial", trControl = train.control)
    
    # Accuracy
    return(model.fit.cv$results[['Accuracy']])
  }
  
  # Fit the model
  glm.fit <- glm(formula =  gene_label ~ ., family = binomial, data = trainingData)

  # predict on test set
  glm.probs <- predict(glm.fit,
                    newdata = testData,
                    type = "response")

  glm.pred <- ifelse(glm.probs > 0.5, 1, 0)

  accuracy <- mean(glm.pred == testData$gene_label)
  print(paste('accuracy:', round(accuracy, 4)))
  
  if (return_probs) {
    return(list(glm.probs, glm.fit, accuracy))
  }
  return(accuracy)
}
```

**LASSO**
```{r}
do_lasso <- function(trainingData, testData, return_probs=FALSE, cv=CROSS_VALIDATE, k=3) {
  if (cv) {
    # find the optimal number for lambda
    x.train <- model.matrix(gene_label~., trainingData)[,-1]
    y.train <- trainingData['gene_label'][[1]]
  
    # Find the best lambda using cross-validation
    cv.lasso <- cv.glmnet(x.train, y.train, alpha = 1, family = "binomial")
    lambda <- cv.lasso$lambda.min
    
    X <- rbind(trainingData[, -which(names(trainingData) == "gene_label")],  testData[, -which(names(testData) == "gene_label")])
    Y <- as.factor(c(trainingData$gene_label, testData$gene_label))
    train.control <- trainControl(method = "cv", number=k)
    
    model.fit.cv <- train(X, Y, method = "glmnet", trControl = train.control, family = 'binomial', tuneGrid = expand.grid(alpha = 1, lambda = lambda))
    
    # Accuracy
    return(model.fit.cv$results[['Accuracy']])
  }

  x.train <- model.matrix(gene_label~., trainingData)[,-1]
  y.train <- trainingData['gene_label'][[1]]
  
  # Find the best lambda using cross-validation
  cv.lasso <- cv.glmnet(x.train, y.train, alpha = 1, family = "binomial")
  print(paste('cv lambda:', cv.lasso$lambda.min))
  
  # Fit the final model on the training data
  model <- glmnet(x.train, y.train, alpha = 1, family = "binomial",
                  lambda = cv.lasso$lambda.min)
  
  # Display regression coefficients
  # coef(model)
  
  # Make predictions on the test data
  x.test <- model.matrix(gene_label ~., testData)[,-1]
  probabilities <- model %>% predict(newx = x.test)
  predicted.classes <- ifelse(probabilities > 0.5, 1, 0)
  
  # Model accuracy
  observed.classes <- testData$gene_label
  accuracy <- mean(predicted.classes == observed.classes)
  print(paste('accuracy:', accuracy))
  if (return_probs) {
    return(list(probabilities, model, accuracy))
  }
  return(accuracy)
}
```

**Random Forest**
```{r}
do_rf <- function(trainingData, testData, return_probs=FALSE, cv=CROSS_VALIDATE, k=3) {
  if (cv) {
    X <- rbind(trainingData[, -which(names(trainingData) == "gene_label")],  testData[, -which(names(testData) == "gene_label")])
    Y <- as.factor(c(trainingData$gene_label, testData$gene_label))
    train.control <- trainControl(method = "cv", number=k)
    tunegrid <- expand.grid(.mtry=sqrt(ncol(X)))
    
    model.fit.cv <- train(X, Y, method = "rf", metric='Accuracy', trControl = train.control, tuneGrid = tunegrid)
    
    # Accuracy
    return(model.fit.cv$results[['Accuracy']])
  }
  
  # trainingData <- transform(trainingData, gene_label=as.factor(gene_label))
  trainingData$gene_label <- as.factor(trainingData$gene_label)
  
  x.train <- model.matrix(gene_label~., trainingData)[,-1]
  y.train <- trainingData['gene_label'][[1]]
  
  x.test <- model.matrix(gene_label~., testData)[,-1]
  y.test <- testData['gene_label'][[1]]

  rf <- randomForest(
    x = x.train,
    y = as.factor(y.train)) #,
    # ntree = 1000)
    #mtry = 25)
  
  # predict on test
  # pred <- predict(rf, newdata=testData[-(dim(testData)[2])])
  prob <- predict(rf, x.test, type="prob")[, 2]
  pred <- ifelse(prob > .5, 1, 0)
  
  # test accuracy
  accuracy <- mean(pred == testData$gene_label)
  
  print(paste('accuracy:', accuracy))
  
  if (return_probs) {
    return(list(prob, rf, accuracy))
  }
  return(accuracy)
}

```


**Gradient Boosting**
```{r warning=FALSE}
do_gb <- function(trainingData, testData, return_probs=FALSE, cv=CROSS_VALIDATE, k=3) {
  if (cv) {
    gb <- gbm(
    formula = gene_label ~ .,
    distribution = 'bernoulli',
    data = trainingData,
    n.trees = 100, 
    cv.folds = 5)
    
    ntree_opt_cv <- gbm.perf(gb, method = 'cv')
    print(paste('optimal number of trees (cv);', ntree_opt_cv))
    
    X <- rbind(trainingData[, -which(names(trainingData) == "gene_label")],  testData[, -which(names(testData) == "gene_label")])
    Y <- as.factor(c(trainingData$gene_label, testData$gene_label))
    train.control <- trainControl(method = "cv", number=k)
    
    gbmGrid <-  expand.grid(n.trees = ntree_opt_cv, interaction.depth=1, shrinkage=.1, n.minobsinnode=10)
    
    model.fit.cv <- train(X, Y, method = "gbm", distribution = 'bernoulli', trControl = train.control, tuneGrid = gbmGrid)

    # Accuracy
    return(model.fit.cv$results[['Accuracy']])
  }
  
  gb <- gbm(
  formula = gene_label ~ .,
  distribution = 'bernoulli',
  data = trainingData,
  n.trees = 100, 
  cv.folds = 5)
  
  ntree_opt_cv <- gbm.perf(gb, method = 'cv')
  print(paste('optimal number of trees (cv);', ntree_opt_cv))
  
  if (FALSE) {
    gb <- gbm(
      formula = gene_label ~ .,
      distribution = 'bernoulli',
      data = trainingData,
      n.trees = ntree_opt_cv, 
      cv.folds = 5)
  }
  
  #Generating a Prediction matrix for each Tree
  probs <-predict(
  object = gb,
  newdata = testData[, -dim(testData)[2]],
  n.trees = ntree_opt_cv,
  type = 'response')

  predictions <- ifelse(probs >= .5, 1, 0)
  accuracy <- mean(predictions == testData$gene_label)
  
  print(paste('accuray:', accuracy))
  
  if (return_probs) {
    return(list(probs, gb, accuracy))
  }
  return(accuracy)
}
```

Make inputData
```{r}
make_inputData <- function(data_df, organ) {
  if (ORGAN_SPECIFIC) {
  inputData <- make_classifier_df(data_df = data_df, organ = , organ, convert_to_rank_perc = CONVERT_TO_RANK_PERC)
  inputData_expr_only <- make_classifier_df(data_df = data_df, organ = organ, convert_to_rank_perc = CONVERT_TO_RANK_PERC, expr_only = TRUE)
} else {
  inputData <- make_classifier_df(data_df = data_df, organ = NULL, convert_to_rank_perc = CONVERT_TO_RANK_PERC)
  inputData <- make_classifier_df(data_df = data_df, organ = NULL, convert_to_rank_perc = CONVERT_TO_RANK_PERC, expr_only = TRUE)
}
  return(list(inputData, inputData_expr_only))
}
```

Train classifiers
```{r}
train_classifiers <- function(data_df, dataset=NULL, organ='brain', return_probs=FALSE, cv=CROSS_VALIDATE) {
  # classifier_df_cur <- data.frame(matrix(ncol=4,nrow=0, dimnames=list(NULL, c('dataset', 'classifier', 'accuracy', 'baseline_accuracy'))))
  
  # split into train and test
  input <- make_inputData(data_df = data_df, organ = organ)
  inputData <- input[[1]]
  inputData_expr_only <- input[[2]]
  
  # split into train and test
  data_train_test <- create_train_test_data(inputData, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
  trainingData <- data_train_test[[1]]
  testData <- data_train_test[[2]]
  
  data_train_test_exp <- create_train_test_data(inputData_expr_only, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
  trainingData_exp <- data_train_test_exp[[1]]
  testData_exp <- data_train_test_exp[[2]]

  # fit all classifiers
  accuracy_logistic <- do_logistic(trainingData, testData)
  accuracy_lasso <- do_lasso(trainingData, testData)
  accuracy_rf <- do_rf(trainingData, testData)
  accuracy_gb <- do_gb(trainingData, testData)
  
  accuracy_logistic_expr_only <- do_logistic(trainingData_exp, testData_exp)
  accuracy_lasso_expr_only <- do_lasso(trainingData_exp, testData_exp)
  accuracy_rf_expr_only <- do_rf(trainingData_exp, testData_exp)
  accuracy_gb_expr_only <- do_gb(trainingData_exp, testData_exp)

  classifier_df <- data.frame(dataset=rep(dataset, 4), classifier=c('Logistic regression', 'LASSO regression', 'Random forest', 'Gradient boosting'), accuracy=c(accuracy_logistic, accuracy_lasso, accuracy_rf, accuracy_gb), baseline_accuracy=c(accuracy_logistic_expr_only, accuracy_lasso_expr_only, accuracy_rf_expr_only, accuracy_gb_expr_only), genes_num=rep(dim(trainingData)[1] + dim(testData)[1], 4))
  
  if (return_probs) {
    # fit all classifiers
    probs_logistic <- do_logistic(trainingData, testData, return_probs=TRUE)
    probs_lasso <- do_lasso(trainingData, testData, return_probs=TRUE)
    probs_rf <- do_rf(trainingData, testData, return_probs=TRUE)
    probs_gb <- do_gb(trainingData, testData, return_probs=TRUE)
    
    probs_logistic_expr_only <- do_logistic(trainingData_exp, testData_exp, return_probs=TRUE)
    probs_lasso_expr_only <- do_lasso(trainingData_exp, testData_exp, return_probs=TRUE)
    probs_rf_expr_only <- do_rf(trainingData_exp, testData_exp, return_probs=TRUE)
    probs_gb_expr_only <- do_gb(trainingData_exp, testData_exp, return_probs=TRUE)
    
    return(list(classifier_df, list(testData$gene_label, probs_logistic, probs_lasso, probs_rf, probs_gb), list(testData_exp$gene_label, probs_logistic_expr_only, probs_lasso_expr_only, probs_rf_expr_only, probs_gb_expr_only)))
  }
    # return classifier_df_cur
    return(list(classifier_df))
}
```

Create the data frame to store the results
```{r}
classifier_df <- data.frame(matrix(ncol=5,nrow=0, dimnames=list(NULL, c('dataset', 'classifier', 'accuracy', 'baseline_accuracy', 'genes_num')))) #, '# genes in each category (risk and non-risk)'))))
```


**1. Tabula Muris (FACS) data**

**1.1 TM Brain (Myeloid) data**
```{r}
tm_brain_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names=1)

dataset <- 'tm_brain_myeloid'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_tm_brain_myeloid <- train_classifiers(data_df = tm_brain_myeloid_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_tm_brain_myeloid[[1]])
```

**1.2 TM Brain (Non-Myeloid) data**
```{r}
# Load filtered and normalized data
tm_brain_non_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Non-Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'tm_brain_non_myeloid'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_tm_brain_non_myeloid <- train_classifiers(data_df = tm_brain_non_myeloid_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_tm_brain_non_myeloid[[1]])
```

**1.3. Tabula Muris(FACS) Heart data.**
```{r}
tm_heart_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Heart-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'tm_heart'
organ <- 'heart'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_tm_heart <- train_classifiers(data_df = tm_heart_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_tm_heart[[1]])
```

**1.4. Tabula Muris(FACS) Lung data.**
# we don't have enough lung genes, so skipping
```{r}
if (FALSE) {
tm_lung_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Lung-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'tm_lung'
organ <- 'lung'

classifier_df_tm_lung <- train_classifiers(data_df = tm_lung_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_tm_lung[[1]])
}
```

**1.5. Tabula Muris(FACS) Kidney data.**
```{r}
tm_kidney_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Kidney-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'tm_kidney'
organ <- 'kidney'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_tm_kidney <- train_classifiers(data_df = tm_kidney_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_tm_kidney[[1]])
```

**2. Zhong 2018 data**
```{r}
zhong2018_data_qc_norm <- read.csv('./scRNA-seq/zhong2018/qc_norm/GSE104276_all_pfc_2394_UMI_count_NOERCC_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'zhong_2018'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_zhong_2018 <- train_classifiers(data_df = zhong2018_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_zhong_2018[[1]])
```

**3. La Manno 2016**
```{r}
lamanno2016_data_qc_norm <- read.csv('./scRNA-seq/lamanno2016/qc_norm/GSE76381_EmbryoMoleculeCounts.cef_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'lamanno_2016'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_lamanno_2016 <- train_classifiers(data_df = lamanno2016_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_lamanno_2016[[1]])
```

**4. Hodge 2019**
```{r}
hodge2019_data_qc_norm <- read.csv('./scRNA-seq/hodge2019/data2/qc_norm/trimmed_means_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'hodge_2019'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_hodge_2019 <- train_classifiers(data_df = hodge2019_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_hodge_2019[[1]])
```

**5. Asp 2019**
```{r}
asp2019_data_qc_norm <- read.csv('./scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm/fixed_all_cells_meta_data_filtered_qc_norm.csv', check.names=FALSE, row.names = 1)

dataset <- 'asp_2019'
organ <- 'heart'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_asp_2019 <- train_classifiers(data_df = asp2019_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_asp_2019[[1]])
```

**6. Nowakowski 2017**
```{r}
nowakowski2017_data_qc_norm <- read.csv('./scRNA-seq/nowakowski2017/qc_norm/exprMatrix_qc_norm.csv', check.names=FALSE, row.names = 1)

rownames(nowakowski2017_data_qc_norm) <- make.unique(tolower(rownames(nowakowski2017_data_qc_norm)))

dataset <- 'nowakowski_2017'
organ <- 'brain'
```

Fit the classifiers and add the results to classifier_df
```{r}
classifier_df_nowakowski_2017 <- train_classifiers(data_df = nowakowski2017_data_qc_norm, dataset = dataset, organ = organ)
classifier_df <- rbind(classifier_df, classifier_df_nowakowski_2017[[1]])
```

Save to disk
```{r}
rownames(classifier_df)=NULL
write.csv(classifier_df, '../results/classifiers/classifier_accuracies_6.csv', row.names = FALSE)
```


**ROC Curves**

Function to plot the ROC curve from a ROC object
```{r}
plot_roc <- function(roc_obj, plot_title='ROC curve') {
  roc_df <- data.frame(
  TPR=rev(roc_obj$sensitivities), 
  FPR=rev(1 - roc_obj$specificities))
  
  ggplot(data=roc_df, aes(x=FPR, y=TPR)) +
    geom_point() +
    geom_line() + 
    geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), col='red', lty=2) +
    xlab('1 - specificity') +
    ylab('sensitivity') +
    ggtitle(plot_title)
}
```

**TM Heart Random forest**
```{r}
input <- make_inputData(data_df = tm_heart_data_qc_norm, organ = 'heart')
inputData <- input[[1]]
inputData_expr_only <- input[[2]]

# split into train and test
data_train_test <- create_train_test_data(inputData, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
trainingData <- data_train_test[[1]]
testData <- data_train_test[[2]]

data_train_test_exp <- create_train_test_data(inputData_expr_only, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
trainingData_exp <- data_train_test_exp[[1]]
testData_exp <- data_train_test_exp[[2]]
```

Fit the classifier
```{r}
# fit rf on the total dataset
tm_heart_rf <- do_rf(trainingData, testData, return_probs=TRUE, cv=FALSE)
tm_heart_rf_exp <- do_rf(trainingData_exp, testData_exp, return_probs=TRUE, cv=FALSE)
```

Create roc objects
```{r}
tm_heart_rf_roc_obj <- roc(testData$gene_label, tm_heart_rf[[1]])
tm_heart_rf_exp_roc_obj <- roc(testData_exp$gene_label, tm_heart_rf_exp[[1]])
```

Plot
```{r}
plot_roc(tm_heart_rf_roc_obj, plot_title = paste('ROC curve for TM Heart data; Random Forest; AUC:', as.character(round(auc(tm_heart_rf_roc_obj), 4))))
```

Plot for baseline
```{r}
plot_roc(tm_heart_rf_exp_roc_obj, plot_title = paste('ROC curve for TM Heart data; Random Forest (baseline); AUC:', as.character(round(auc(tm_heart_rf_exp_roc_obj), 4))))
```

**Feature Importance**

```{r}
fi_tm_heart <- data.frame(tm_heart_rf[[2]]$importance)
fi_tm_heart$feature <- rownames(fi_tm_heart)
fi_tm_heart <- fi_tm_heart[order(fi_tm_heart$MeanDecreaseGini, decreasing = TRUE),]
View(fi_tm_heart)
```


**TM Brain Non-Myeloid**
```{r}
input <- make_inputData(data_df = tm_brain_non_myeloid_data_qc_norm, organ = 'heart')
inputData <- input[[1]]
inputData_expr_only <- input[[2]]

# split into train and test
data_train_test <- create_train_test_data(inputData, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
trainingData <- data_train_test[[1]]
testData <- data_train_test[[2]]

data_train_test_exp <- create_train_test_data(inputData_expr_only, balanced_test = BALANCED_TEST, balanced_train = BALANCED_TRAIN)
trainingData_exp <- data_train_test_exp[[1]]
testData_exp <- data_train_test_exp[[2]]
```

Fit the classifier
```{r}
# fit rf on the total dataset
tm_brain_non_myeloid_rf <- do_rf(trainingData, testData, return_probs=TRUE, cv=FALSE)
tm_brain_non_myeloid_rf_exp <- do_rf(trainingData_exp, testData_exp, return_probs=TRUE, cv=FALSE)
```

Create roc objects
```{r}
tm_brain_non_myeloid_rf_roc_obj <- roc(testData$gene_label, tm_brain_non_myeloid_rf[[1]])
tm_brain_non_myeloid_rf_exp_roc_obj <- roc(testData_exp$gene_label, tm_brain_non_myeloid_rf_exp[[1]])
```

Plot
```{r}
plot_roc(tm_brain_non_myeloid_rf_roc_obj, plot_title = paste('ROC curve for TM Brain (Non-Myeloid) data; Random Forest; AUC:', as.character(round(auc(tm_brain_non_myeloid_rf_roc_obj), 4))))
```

Plot for baseline
```{r}
plot_roc(tm_brain_non_myeloid_rf_exp_roc_obj, plot_title = paste('ROC curve for TM Brain (Non-Myeloid) data; Random Forest (baseline);\nAUC:', as.character(round(auc(tm_brain_non_myeloid_rf_exp_roc_obj), 4))))
```


**Feature Importance**

```{r}
fi_tm_nonmyeloid <- data.frame(tm_brain_non_myeloid_rf[[2]]$importance)
fi_tm_nonmyeloid$feature <- rownames(fi_tm_nonmyeloid)
fi_tm_nonmyeloid <- fi_tm_nonmyeloid[order(fi_tm_nonmyeloid$MeanDecreaseGini, decreasing = TRUE),]
View(fi_tm_nonmyeloid)
```