---
title: "Fisher's Exact Test"
author: "Ani Khachatryan"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='E:/Columbia/Spring2020/lab/data/')
```

Load the risk genes.
```{r}
DDD_all <- read.csv('./DDD/DDG2P_28_1_2020.csv')

DDD_all$organ.specificity.list <- tolower(DDD_all$organ.specificity.list)
DDD_all$gene.symbol <- tolower(DDD_all$gene.symbol)

DDD <- subset(DDD_all, DDD.category == 'confirmed')
```

Create a T/F table of genes vs RBPs, TRUE if RBP regulates the gene
rbp_threshold is the number of binding sites/events (?) necessary to consider the RBP a regulator of the protein

```{r}
# RBP-regualted genes
rbp_table <- read.table('./RBP_targets/RBP_targets.txt', header=TRUE, row.names = 1)
colnames(rbp_table) <- tolower(colnames(rbp_table))
rownames(rbp_table) <- tolower(rownames(rbp_table))
rbps_w_celltypes <- colnames(rbp_table)

sfun <- function(MyString) {
  return(strsplit(MyString, '.', TRUE)[[1]][2])
}

rbps <- sapply(rbps_w_celltypes, sfun)
#rbps <- tolower(sapply(rbps_w_celltypes, sfun))
colnames(rbp_table) <- rbps

# needs to have at least rbp_threshold binding sites 
rbp_threshold <- 10
rbp_table_tf <- rbp_table >= rbp_threshold
```

Fisher's exact test.
```{r}
fishers_exact_test <- function(data_df, celltype, organ, rbp_quantile_threshold=.9, gene_quanitle_threshold=.5, consider_organ=TRUE) {
  rownames(data_df) <- make.unique(tolower(rownames(data_df)))
  print(paste('cell type:', celltype))
  
  # some preprocessing
  # df_ct is the subset of the passed df, but only includes columns of the given celltype
  df_ct <- data_df[, grep(celltype, names(data_df))]
  # then we take the mean of all the cells to get the mean expression values for the celltype
  df_ct_means <- rowMeans(df_ct)
  # only keep top 50% expressed genes
  quantile_gene <- quantile(df_ct_means, gene_quanitle_threshold)
  df_ct_means_top50<- df_ct_means[df_ct_means >= quantile_gene]

  # find important RBP threshhold in celltype 
  quantile_rbp <- quantile(df_ct_means, rbp_quantile_threshold)
  
  # finding important RBPs in this cell type
  # and RBP is considered important for this cell type if it's among the top 10% expressed genes
  important_rbps_ct <- names(df_ct_means_top50[rbps][df_ct_means_top50[rbps] >= quantile_rbp])
  important_rbps_ct <- na.omit(important_rbps_ct)
  
  # filtering rbp_table_tf to only keep important RBPs for this cell type
  rbp_table_tf_important_rbps <- rbp_table_tf[, important_rbps_ct]

  # separate genes into risk and non-riks
  # remove non-confirmed DDD risk genes from non-risk
  # RISK GENES  
  if (consider_organ) {
    risk_genes <- unique(DDD[grep(organ, DDD$organ.specificity.list, ignore.case = TRUE),]$gene.symbol)
  } else {
    risk_genes <- unique(DDD$gene.symbol)
  }
  
  risk_genes <- names(df_ct_means_top50)[names(df_ct_means_top50) %in% risk_genes]
  non_risk_genes <- names(df_ct_means_top50)[!(names(df_ct_means_top50) %in% risk_genes)]
  non_risk_genes <- non_risk_genes[!(non_risk_genes %in% DDD_all$gene.symbol)]
  
  # number of top 50% expressed risk genes
  n <- length(risk_genes)
  
  # number of top 50% expressed non-risk genes
  m <- length(non_risk_genes)
  
  # total number of genes in consideration
  N <- n + m
  
  # number of risk genes regulated by important RBPs
  a <- sum(risk_genes %in% names(genes_regulated_by_important_rbps_ct))
  b <- n - a
  
  # number of non-risk genes regulated by important RBPs
  c <- sum(non_risk_genes %in% names(genes_regulated_by_important_rbps_ct))
  d <- m - c
  
  # create the table for Fisher's extact test
  fishers_table <- matrix(data = c(a,b,c,d), nrow = 2, ncol = 2, dimnames = list( c('RBP', 'non-RBP'), c('risk', 'non-risk')))
  
  print(fishers_table)
  ft_result <- fisher.test(fishers_table)
  
  print(paste0('p-value: ', ft_result$p.value))
  print(paste0('odds ratio: ', ft_result$estimate))
  
  return(ft_result)
  
}
```

Helper function: removes the extension from the cell type
```{r}
remove_ext <- function(mystr) {
    return(tools::file_path_sans_ext(mystr))
  }
```

Create a data.frame to store the results
```{r}
ft_df <- data.frame(matrix(ncol=4,nrow=0, dimnames=list(NULL, c('dataset', 'celltype', 'pval', 'odds_ratio'))))
```

**1. Tabula Muris (FACS) data**

**1.1 TM Brain (Myeloid) data**
```{r}
tm_brain_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names=1)

colnames(tm_brain_myeloid_data_qc_norm) <- unlist(lapply(colnames(tm_brain_myeloid_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'tm_brain_myeloid'
celltypes <- unique(colnames(tm_brain_myeloid_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = tm_brain_myeloid_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}

```


**1.2 TM Brain (Non-Myeloid) data**
```{r}
# Load filtered and normalized data
tm_brain_non_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Non-Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(tm_brain_non_myeloid_data_qc_norm) <- unlist(lapply(colnames(tm_brain_non_myeloid_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'tm_brain_non_myeloid'
celltypes <- unique(colnames(tm_brain_non_myeloid_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = tm_brain_non_myeloid_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```


**1.3. Tabula Muris(FACS) Heart data.**
```{r}
tm_heart_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Heart-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(tm_heart_data_qc_norm) <- unlist(lapply(colnames(tm_heart_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'tm_heart'
celltypes <- unique(colnames(tm_heart_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = tm_heart_data_qc_norm, celltype = current_celltype, organ = 'heart')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```


**1.4. Tabula Muris(FACS) Lung data.**
```{r}
tm_lung_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Lung-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(tm_lung_data_qc_norm) <- unlist(lapply(colnames(tm_lung_data_qc_norm), remove_ext))
```


Fisher's exact test
```{r}
dataset <- 'tm_lung'
celltypes <- unique(colnames(tm_lung_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = tm_lung_data_qc_norm, celltype = current_celltype, organ = 'lung')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```

**1.5. Tabula Muris(FACS) Kidney data.**
```{r}
tm_kidney_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Kidney-counts_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(tm_kidney_data_qc_norm) <- unlist(lapply(colnames(tm_kidney_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'tm_kidney'
celltypes <- unique(colnames(tm_kidney_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = tm_kidney_data_qc_norm, celltype = current_celltype, organ = 'kidney')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```

**2. Zhong 2018 data**
```{r}
zhong2018_data_qc_norm <- read.csv('./scRNA-seq/zhong2018/qc_norm/GSE104276_all_pfc_2394_UMI_count_NOERCC_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(zhong2018_data_qc_norm) <- unlist(lapply(colnames(zhong2018_data_qc_norm), remove_ext))

```

Fisher's exact test
```{r}
dataset <- 'zhong_2018'
celltypes <- unique(colnames(zhong2018_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = zhong2018_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```


**3. La Manno 2016**
```{r}
lamanno2016_data_qc_norm <- read.csv('./scRNA-seq/lamanno2016/qc_norm/GSE76381_EmbryoMoleculeCounts.cef_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(lamanno2016_data_qc_norm) <- unlist(lapply(colnames(lamanno2016_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'lamanno_2016'
celltypes <- unique(colnames(lamanno2016_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = lamanno2016_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```

**4. Hodge 2019**
```{r}
hodge2019_data_qc_norm <- read.csv('./scRNA-seq/hodge2019/data2/qc_norm/trimmed_means_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(hodge2019_data_qc_norm) <- unlist(lapply(colnames(hodge2019_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'hodge_2019'
celltypes <- unique(colnames(hodge2019_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = hodge2019_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```

**5. Asp 2019**
```{r}
asp2019_data_qc_norm <- read.csv('./scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm/fixed_all_cells_meta_data_filtered_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(asp2019_data_qc_norm) <- unlist(lapply(colnames(asp2019_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'asp_2019'
celltypes <- unique(colnames(asp2019_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = asp2019_data_qc_norm, celltype = current_celltype, organ = 'heart')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```


**6. Nowakowski 2017**
```{r}
nowakowski2017_data_qc_norm <- read.csv('./scRNA-seq/nowakowski2017/qc_norm/exprMatrix_qc_norm.csv', check.names=FALSE, row.names = 1)

colnames(nowakowski2017_data_qc_norm) <- unlist(lapply(colnames(nowakowski2017_data_qc_norm), remove_ext))
```

Fisher's exact test
```{r}
dataset <- 'nowakowski_2017'
celltypes <- unique(colnames(nowakowski2017_data_qc_norm))

for (current_celltype in celltypes) {
  ft_result <- fishers_exact_test(data_df = nowakowski2017_data_qc_norm, celltype = current_celltype, organ = 'brain')

  ft_df <- rbind(ft_df, data.frame(dataset=dataset, celltype=current_celltype, pval=ft_result$p.value, odds_ratio=ft_result$estimate))
}
```


```{r}
rownames(ft_df)=NULL
write.csv(ft_df, '../results/fishers_exact_test/fishers_test.csv', row.names = FALSE)
```

