---
title: "PCA"
author: "Ani Khachatryan"
date: "3/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='E:/Columbia/Spring2020/lab/data/scRNA-seq/')
```

Imports.
```{r}
library(ggplot2)
library(tools)
```


Let's define the function that will take a the count matrix (data.frame), perform PCA on it, and plot the results.
```{r}
pca_and_plot <- function(data_df, plot_title, plot_path, remove_outliers=TRUE, colnames_are_celltypes=FALSE) {
  myfun <- function(mystr) {
    return(tools::file_path_sans_ext(mystr))
  }
  
  # make sure all columns are numeric
  data_df <- sapply(data_df, as.numeric)
  
  # remove 0 expression columns
  data_df[,colSums(data_df) != 0]

  # PCA
  pca <- prcomp(t(as.matrix(data_df)), scale=TRUE) 
  
  pca.var <- pca$sdev^2
  pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
  
  pca.data <- data.frame(Sample=rownames(pca$x),
  X=pca$x[,1],
  Y=pca$x[,2])
  
  # remove outliers
  if (remove_outliers) {
    # X
    # Find the interquartile range of X
    iqr_X <- as.numeric(quantile(pca.data$X, .75) - quantile(pca.data$X, .25))
    # Find the median of X
    median_X <- as.numeric(median(pca.data$X))
    
    # Remove outliers of X (median +- 3*iqr)
    pca.data <- pca.data[pca.data$X >= median_X - 3*iqr_X,]
    pca.data <- pca.data[pca.data$X <= median_X + 3*iqr_X,]
    
    # Y
    # Find the interquartile range of Y
    iqr_Y <- as.numeric(quantile(pca.data$Y, .75) - quantile(pca.data$Y, .25))
     # Find the median of Y
    median_Y <- as.numeric(median(pca.data$Y))
    
    # Remove outliers of Y (median +- 3*iqr)
    pca.data <- pca.data[pca.data$Y >= median_Y - 3*iqr_Y,]
    pca.data <- pca.data[pca.data$Y <= median_Y + 3*iqr_Y,]
  }
  
  if (!colnames_are_celltypes) {
    pca.data$Sample <- sapply(pca.data$Sample, myfun)
  }
  pca.data$Sample <- as.factor(as.character(pca.data$Sample))
  
  # p <- 
  ggplot(data=pca.data, aes(x=X, y=Y, color=Sample)) +
  geom_point() +
  xlab(paste("PC1 - ", pca.var.per[1], "%", sep="")) +
  ylab(paste("PC2 - ", pca.var.per[2], "%", sep="")) +
  ggtitle(plot_title)
  
  # ggsave(plot=p, filename=gsub(' ', '_', plot_title, fixed = TRUE), path=plot_path)
}
```

Helpful function to replace spaces in colnames with underscores.
```{r}
space_to_underscore <- function(mystr) {
  return(gsub(' ', '_', mystr, fixed=TRUE))
}
```


**Now let's start doing and plotting PCA on each of the scRNA-seq datasets.**

**1. Tabula Muris (FACS) data**
**1.1. Tabula Muris (FACS) Brain(Myeloid) data.**

Normalized.
```{r}
# Load normalized data
tm_brain_myeloid_data_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Brain_Myeloid-counts_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_brain_myeloid_data_norm, plot_title='Tabula Muris Brain (Myeloid) data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE)

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
tm_brain_myeloid_data_qc_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Brain_Myeloid-counts_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_brain_myeloid_data_qc_norm, plot_title='Tabula Muris Brain (Myeloid) data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE)
```

**1.2. Tabula Muris(FACS) Brain (Non-myeloid) data.**

Normalized.
```{r}
# Load normalized data
tm_brain_non_myeloid_data_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Brain_Non-Myeloid-counts_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_brain_non_myeloid_data_norm, plot_title='Tabula Muris Brain (Non-Myeloid) data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE)

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
tm_brain_non_myeloid_data_qc_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Brain_Non-Myeloid-counts_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_brain_non_myeloid_data_qc_norm, plot_title='Tabula Muris Brain (Non-Myeloid) data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE)
```

**1.3. Tabula Muris(FACS) Heart data.**

Normalized.
```{r}
# Load normalized data
tm_heart_data_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Heart-counts_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_heart_data_norm, plot_title='Tabula Muris Heart data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
tm_heart_data_qc_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Heart-counts_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_heart_data_qc_norm, plot_title='Tabula Muris Heart data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**1.4. Tabula Muris(FACS) Lung data.**

Normalized.
```{r}
# Load normalized data
tm_lung_data_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Lung-counts_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_lung_data_norm, plot_title='Tabula Muris Lung data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
tm_lung_data_qc_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Lung-counts_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_lung_data_qc_norm, plot_title='Tabula Muris Lung data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**1.5. Tabula Muris(FACS) Kidney data.**

Normalized.
```{r}
# Load normalized data
tm_kidney_data_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Kidney-counts_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_kidney_data_norm, plot_title='Tabula Muris Kidney data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
tm_kidney_data_qc_norm <- read.csv('./TabulaMuris/FACS/qc_norm/Kidney-counts_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=tm_kidney_data_qc_norm, plot_title='Tabula Muris Kidney data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**2. Zhong 2018 data**

Normalized.
```{r}
# Load normalized data
zhong2018_data_norm <- read.csv('./zhong2018/qc_norm/GSE104276_all_pfc_2394_UMI_count_NOERCC_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=zhong2018_data_norm, plot_title='Zhong 2018 data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
zhong2018_data_qc_norm <- read.csv('./zhong2018/qc_norm/GSE104276_all_pfc_2394_UMI_count_NOERCC_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=zhong2018_data_qc_norm, plot_title='Zhong 2018 data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**3. La Manno 2016**

Normalized.
```{r}
# Load normalized data
lamanno2016_data_norm <- read.csv('./lamanno2016/qc_norm/GSE76381_EmbryoMoleculeCounts.cef_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=lamanno2016_data_norm, plot_title='La Manno 2016 data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
lamanno2016_data_qc_norm <- read.csv('./lamanno2016/qc_norm/GSE76381_EmbryoMoleculeCounts.cef_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=lamanno2016_data_qc_norm, plot_title='La Manno 2016 data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**4. Hodge 2019**

Normalized.
```{r}
# Load normalized data
hodge2019_data_norm <- read.csv('./hodge2019/data2/qc_norm/trimmed_means_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=hodge2019_data_norm, plot_title='Hodge 2019 data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
hodge2019_data_qc_norm <- read.csv('./hodge2019/data2/qc_norm/trimmed_means_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=hodge2019_data_qc_norm, plot_title='Hodge 2019 data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**5. Asp 2019**

Normalized.
```{r}
# Load normalized data
asp2019_data_norm <- read.csv('./asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm/fixed_all_cells_meta_data_filtered_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=asp2019_data_norm, plot_title='Asp 2019 data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
asp2019_data_qc_norm <- read.csv('./asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm/fixed_all_cells_meta_data_filtered_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=asp2019_data_qc_norm, plot_title='Asp 2019 data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```

**6. Nowakowski 2017**
Normalized.
```{r}
# Load normalized data
nowakowski2017_data_norm <- read.csv('./nowakowski2017/qc_norm/exprMatrix_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=nowakowski2017_data_norm, plot_title='Nowakowski 2017 data post normalization', plot_path='../../results/pca/', remove_outliers=TRUE )

```

Filtered & normalized.
```{r}
# Load filtered and normalized data
nowakowski2017_data_qc_norm <- read.csv('./nowakowski2017/qc_norm/exprMatrix_qc_norm.csv', check.names=FALSE)

# Do PCA
pca_and_plot(data_df=nowakowski2017_data_qc_norm, plot_title='Nowakowski 2017 data post QC and normalization', plot_path='../../results/pca/', remove_outliers=TRUE )
```


