---
title: "QC and Normalization"
author: "Ani Khachatryan"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='E:/Columbia/Spring2020/lab')
```

Imports.

```{r}
library(Seurat)
library(tidyr)
```

Quality Control and Normalization for scRNA-seq data

**First of all, let's define the function that will take a dataset and perform astandard QC and log-normalization on it **

QC and normalization (SCTransform) function.
```{r}
# FUNCTION: qc_and_normalize
# ARGUMENTS: 
  # data_df - the count matrix as a data.frame object
  # coeff - for thresholding
    #default: 3
  # normalize_only - will normalize without QC
# RETURNS: 
  # a data.frame (filtered and normalized)

qc_and_normalize_2 <- function(data_df, coeff=3, normalize_only=FALSE) {
  if (!normalize_only) {
    ncols_before <- dim(data_df)[2]
  
    # filter based on the number of expressed genes per sample (high and low)
    features_RNA <- colSums(data_df > 0)
    nFeature_RNA_upper <- median(features_RNA) + coeff*mad(features_RNA)
    nFeature_RNA_lower <- median(features_RNA) - coeff*mad(features_RNA)
    
    # filter the count matrix
    data_df_qc <- data_df[,features_RNA > nFeature_RNA_lower & features_RNA < nFeature_RNA_upper]
    
    #filter based on the total RNA counts per sample (high and low)
    # counts_RNA <- colSums(data_df_qc)
    counts_RNA <- colSums(data_df)
    
    nCount_RNA_upper <- median(counts_RNA) + coeff*mad(counts_RNA)
    nCount_RNA_lower <- median(counts_RNA) - coeff*mad(counts_RNA)
    
    # filter the count matrix
    data_df_qc <- data_df_qc[,counts_RNA > nCount_RNA_lower & counts_RNA < nCount_RNA_upper]
    
    ncols_after <- dim(data_df_qc)[2]
    
    perc <- (ncols_before-ncols_after)/ncols_before * 100
    # print(paste0('removed ', round(perc,2), '% of samples'))
  } else {
      data_df_qc <- data_df
  }
  
  data_seurat <- CreateSeuratObject(counts = data_df_qc)
  data_seurat <- SCTransform(data_seurat, verbose = FALSE)
  
  return(as.data.frame(data_seurat@assays$SCT@counts))
}
```

Define a function that will take the data, directory, and file name and write the normalized and qc-normalized .csv files to the directory
```{r}
qc_norm_write <- function(data_df, directory, file_name) {
  setwd(directory)
  
  # normalize without QC.
  data_df_norm <- qc_and_normalize_2(data_df=data_df, coeff=3, normalize_only=TRUE)
  # write to disc
  write.csv(data_df_norm, paste0(file_name, '_norm.csv'))

  # Perform QC and normalization.
  data_df_qc_norm <- qc_and_normalize_2(data_df=data_df)
  # write to disc
  write.csv(data_df_qc_norm, paste0(file_name, '_qc_norm.csv'))
}
```

Define a function that will replace sample_ids with cell_types for the data
```{r}
id_to_celltype <- function(data_df, annotation_df) {
  myfun <- function(sample_id) {
    return(as.character(annotation_df$cell_type[annotation_df$sample_id == as.character(sample_id)])[1])
  }
  colnames(data_df) <- sapply(colnames(data_df), myfun)
  
  # replace NA values with 'unknown'
  colnames(data_df) <- tidyr::replace_na(colnames(data_df), 'unknown')
  
  # replace empty strings with 'unknown'
  colnames(data_df)[colnames(data_df) == ''] <- 'unknown'
  
  return(data_df)
}
```

**Now let's start doing QC and normalization on each of the scRNA-seq datasets.**
We will also reaname the column names to cell types, where applicable.

**1. Tabula Muris (FACS) data - annotations.**
```{r}
tm_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/qc_norm'

tm_annotations <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/annotations_facs.csv')

# we need to name the necessary columns to sample_id and cell_type
colnames(tm_annotations)[3] <- 'sample_id'
colnames(tm_annotations)[4] <- 'cell_type'
```

**1.1. Tabula Muris (FACS) Brain(Myeloid) data.**
```{r warning=FALSE}
# Brain (Myeloid)
# Load the data
tm_brain_myeloid_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/Brain_Myeloid-counts.csv', row.names=1)

# rename colnames to be cell types
tm_brain_myeloid_data <- id_to_celltype(data_df=tm_brain_myeloid_data, annotation_df=tm_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=tm_brain_myeloid_data, directory=tm_directory, file_name='Brain_Myeloid-counts')
```

**1.2. Tabula Muris(FACS) Brain (Non-myeloid) data.**
```{r warning=FALSE}
# Brain (Non-myeloid)
# Load the data
tm_brain_non_myeloid_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/Brain_Non-Myeloid-counts.csv', row.names=1)

# rename colnames to be cell types
tm_brain_non_myeloid_data <- id_to_celltype(data_df=tm_brain_non_myeloid_data, annotation_df=tm_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=tm_brain_non_myeloid_data, directory=tm_directory, file_name='Brain_Non-Myeloid-counts')
```

**1.3. Tabula Muris(FACS) Heart data.**
```{r warning=FALSE}
# Heart
# Load the data
tm_heart_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/Heart-counts.csv', row.names=1)

# rename colnames to be cell types
tm_heart_data <- id_to_celltype(data_df=tm_heart_data, annotation_df=tm_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=tm_heart_data, directory=tm_directory, file_name='Heart-counts')
```

**1.4. Tabula Muris(FACS) Lung data.**
```{r warning=FALSE}
# Lung
# Load the data
tm_lung_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/Lung-counts.csv', row.names=1)

# rename colnames to be cell types
tm_lung_data <- id_to_celltype(data_df=tm_lung_data, annotation_df=tm_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=tm_lung_data, directory=tm_directory, file_name='Lung-counts')
```

**1.5. Tabula Muris(FACS) Kidney data.**
```{r warning=FALSE}
# Kidney
# Load the data
tm_kidney_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/TabulaMuris/FACS/Kidney-counts.csv', row.names=1)

# rename colnames to be cell types
tm_kidney_data <- id_to_celltype(data_df=tm_kidney_data, annotation_df=tm_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=tm_kidney_data, directory=tm_directory, file_name='Kidney-counts')
```

**2. Zhong 2018 data**
```{r warning=FALSE}
zhong2018_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/zhong2018/qc_norm'

# Load the annotations
zhong2018_annotations <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/zhong2018/SampleInfo.csv')

colnames(zhong2018_annotations)[1] <- 'sample_id'
colnames(zhong2018_annotations)[2] <- 'cell_type'

# zhong 2018
# I am using the UMI count data, not the TPM data
# Load the data
zhong2018_data <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/zhong2018/GSE104276_all_pfc_2394_UMI_count_NOERCC.xls')
# Remove 0 expression samples
zhong2018_data <- zhong2018_data[,colSums(zhong2018_data) != 0]

# rename colnames to be cell types
zhong2018_data <- id_to_celltype(data_df=zhong2018_data, annotation_df=zhong2018_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=zhong2018_data, directory=zhong2018_directory, file_name='GSE104276_all_pfc_2394_UMI_count_NOERCC')
```

**3. La Manno 2016**
# in this data, the columns are already cell types
```{r warning=FALSE}
# dirctory to write to
lamanno2016_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/lamanno2016/qc_norm'

# Load the data
lamanno2016_data <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/lamanno2016/GSE76381_EmbryoMoleculeCounts.cef.txt', sep ="\t", skip=5, header = TRUE, row.names=1)
lamanno_header <- scan('E:/Columbia/Spring2020/lab/data/scRNA-seq/lamanno2016/GSE76381_EmbryoMoleculeCounts.cef.txt', '', skip = 2, nlines = 1, sep = '\t')
lamanno_header <- lamanno_header[3:length(lamanno_header)]
lamanno2016_data <-lamanno2016_data[ -c(1) ]
colnames(lamanno2016_data) <- lamanno_header


# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=lamanno2016_data, directory=lamanno2016_directory, file_name='GSE76381_EmbryoMoleculeCounts.cef')
```

**4. Hodge 2019**
```{r warning=FALSE}
hodge2019_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/hodge2019/data2/qc_norm'

# Load the annotations
hodge2019_annotations <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/hodge2019/data2/sample_annotations.csv')

colnames(hodge2019_annotations)[4] <- 'sample_id'
colnames(hodge2019_annotations)[7] <- 'cell_type'

# Extra step for Hodge 2019 data.
hodge2019_annotations$sample_id <- gsub(" ", ".", hodge2019_annotations$sample_id)

# Hodge 2019
# Load the data
hodge2019_data <- read.csv('E:/Columbia/Spring2020/lab/data/scRNA-seq/hodge2019/data2/trimmed_means.csv', row.names = 1)

# Extra step for Hodge 2019 data
multiply_factor <- 10000
hodge2019_data <- hodge2019_data * multiply_factor
hodge2019_rownames <- rownames(hodge2019_data)
hodge2019_data <- as.data.frame(sapply(hodge2019_data, as.integer))
rownames(hodge2019_data) <- hodge2019_rownames

# rename colnames to be cell types
hodge2019_data <- id_to_celltype(data_df=hodge2019_data, annotation_df=hodge2019_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=hodge2019_data, directory=hodge2019_directory, file_name='trimmed_means')
```

**5. Asp 2019**
Quick fix for the annotation file.
```{r warning=FALSE}
# fixing the annotation files

# removing extra lines
mytsv <- readLines('E:/Columbia/Spring2020/lab/data/scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/all_cells_meta_data_filtered.tsv')


# the charactes which are allowed to be the first character of a line
allowed_chars <- c('A', 'C', 'T', 'G')

myfun <- function(myline) {
  if (sum(startsWith(myline, allowed_chars))) {
    return(TRUE)
  }
  
  return(FALSE)
}

lines_filter <- lapply(mytsv, myfun)
mytsv_filtered <- mytsv[unlist(lines_filter)]

#### 
# fixing the missing 8th column issue
myfun2 <- function(mystr) {
  if (endsWith(mystr, ' ')) {
    return(paste0(gsub(" ", "", mystr), "\tNA"))
  }
  
  return(mystr)
}

mytsv_filtered_fixed <- unlist(lapply(mytsv_filtered, myfun2))


setwd('E:/Columbia/Spring2020/lab/data/scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files')

write(mytsv[1], file='fixed_all_cells_meta_data_filtered.tsv', append=FALSE, sep='\t')
write(mytsv_filtered_fixed, file='fixed_all_cells_meta_data_filtered.tsv', append=TRUE, sep='\t')
```

```{r warning=FALSE}

asp2019_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm'

# Load the annotations
asp2019_annotations <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/fixed_all_cells_meta_data_filtered.tsv', colClasses = c(NA, NA, NA, NA, NA, NA, NA, NULL), sep='\t', header = TRUE)

colnames(asp2019_annotations)[1] <- 'sample_id'
colnames(asp2019_annotations)[7] <- 'cell_type'

# Asp 2019
# Load the data.
asp2019_data <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/all_cells_count_matrix_filtered.tsv', sep='\t', header = TRUE, row.names = 1)

# rename colnames to be cell types
asp2019_data <- id_to_celltype(data_df=asp2019_data, annotation_df=asp2019_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=asp2019_data, directory=asp2019_directory, file_name='fixed_all_cells_meta_data_filtered')
```

**6. Nowakowski 2017**
# remove outliers
Load the annotations.
```{r warning=FALSE}
nowakowski2017_directory <- 'E:/Columbia/Spring2020/lab/data/scRNA-seq/nowakowski2017/qc_norm'
  
nowakowski2017_annotations <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/nowakowski2017/meta.tsv', sep='\t', header=TRUE)

colnames(nowakowski2017_annotations)[1] <- 'sample_id'
colnames(nowakowski2017_annotations)[2] <- 'cell_type'

# Nowakowski 2017
# Load the data
nowakowski2017_data <- read.table('E:/Columbia/Spring2020/lab/data/scRNA-seq/nowakowski2017/exprMatrix.tsv.gz.tmp', row.names = 1, header=TRUE)

# Extra step for Nowakowski 2017 data.
# multiply_factor <- 1000
# nowakowski2017_data <- nowakowski2017_data * multiply_factor

nowakowski2017_rownames <- rownames(nowakowski2017_data)
nowakowski2017_data <- as.data.frame(sapply(nowakowski2017_data, as.integer))
rownames(nowakowski2017_data) <- nowakowski2017_rownames

# rename colnames to be cell types
nowakowski2017_data <- id_to_celltype(data_df=nowakowski2017_data, annotation_df=nowakowski2017_annotations)

# perform normalization and qc-normalization and write to disk
qc_norm_write(data_df=nowakowski2017_data, directory=nowakowski2017_directory, file_name='exprMatrix')
```