---
title: "UMAP"
author: "Ani Khachatryan"
date: "4/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='E:/Columbia/Spring2020/lab/data/')
```

Imports
```{r}
library(umap)
```

Helper function: removes the extension from the cell type
```{r}
remove_ext <- function(mystr) {
    return(tools::file_path_sans_ext(mystr))
  }
```

Performs UMAP.
```{r}
do_umap <- function(data_df) {
  # genes <- rownames(data_df)
  
  data_df <- data.frame(t(data_df))
  data.labels <- unlist(lapply(rownames(data_df), remove_ext))
  
  # colnames(data_df) <- genes
  # data_df <- data_df[-1,]
  
  data.umap <- umap(data_df)
  data.plot <- cbind(data.umap$layout, data.frame(Sample=data.labels))
  
  return(data.plot)
}
```

Plots the UMAP results.
```{r}
plot_umap <- function(data.plot, plot_title) {
  ggplot(data=data.plot, aes(x=`1`, y=`2`, color=Sample)) +
  geom_point() +
  xlab('x') +
  ylab('y') +
  ggtitle(plot_title)
}
```


**1. Tabula Muris (FACS) data**

**1.1 TM Brain (Myeloid) data**
```{r}
tm_brain_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names=1)
```

Perform UMAP
```{r}
data.plot <- do_umap(tm_brain_myeloid_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Tabula Muris Brain (Myeloid) data post QC and normalization')
```


**1.2 TM Brain (Non-Myeloid) data**
```{r}
# Load filtered and normalized data
tm_brain_non_myeloid_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Brain_Non-Myeloid-counts_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(tm_brain_non_myeloid_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Tabula Muris Brain (Non-Myeloid) data post QC and normalization')
```


**1.3. Tabula Muris(FACS) Heart data.**
```{r}
tm_heart_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Heart-counts_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(tm_heart_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Tabula Muris Heart data post QC and normalization')
```


**1.4. Tabula Muris(FACS) Lung data.**
```{r}
tm_lung_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Lung-counts_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(tm_lung_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Tabula Muris Lung data post QC and normalization')
```


**1.5. Tabula Muris(FACS) Kidney data.**
```{r}
tm_kidney_data_qc_norm <- read.csv('./scRNA-seq/TabulaMuris/FACS/qc_norm/Kidney-counts_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(tm_kidney_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Tabula Muris Kidney data post QC and normalization')
```


**2. Zhong 2018 data**
```{r}
zhong2018_data_qc_norm <- read.csv('./scRNA-seq/zhong2018/qc_norm/GSE104276_all_pfc_2394_UMI_count_NOERCC_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(zhong2018_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Zhong 2018 data post QC and normalization')
```


**3. La Manno 2016**
```{r}
lamanno2016_data_qc_norm <- read.csv('./scRNA-seq/lamanno2016/qc_norm/GSE76381_EmbryoMoleculeCounts.cef_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(lamanno2016_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'La Manno 2016 data post QC and normalization')
```

**4. Hodge 2019**
```{r}
hodge2019_data_qc_norm <- read.csv('./scRNA-seq/hodge2019/data2/qc_norm/trimmed_means_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(hodge2019_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Hodge 2019 data post QC and normalization')
```


**5. Asp 2019**
```{r}
asp2019_data_qc_norm <- read.csv('./scRNA-seq/asp2019/filtered_count_matrices_and_meta_tables/share_files/qc_norm/fixed_all_cells_meta_data_filtered_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(asp2019_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Asp 2019 data post QC and normalization')
```


**6. Nowakowski 2017**
```{r}
nowakowski2017_data_qc_norm <- read.csv('./scRNA-seq/nowakowski2017/qc_norm/exprMatrix_qc_norm.csv', check.names=FALSE, row.names = 1)
```

Perform UMAP
```{r}
data.plot <- do_umap(nowakowski2017_data_qc_norm)
```

Plot the results
```{r}
plot_umap(data.plot, 'Nowakowski 2017 data post QC and normalization')
```

